### 2013-07-15 Rcode/intro.rmd
updated test of vb method; updated intro

Note that $\text{lfsr}_j \geq \text{lfdr}_j$ 
because both the events $\beta_j \geq 0$
and $\beta_j \leq 0$ include the event $\beta_j=0$.
Thus, lfsr gives an upper bound for the lfdr,
and so can be used as a conservative estimate of the lfdr
if an lfdr is desired. (This may be helpful when comparing
with methods that compute an lfdr.)

@@

beta.ash=ash(ss$betahat,ss$betasd)
beta.ash.auto = ash(ss$betahat, ss$betasd,auto=TRUE)
#these to test the VB version 
#beta.ash.vb.uniform = ash(ss$betahat, ss$betasd,auto=TRUE, VB=TRUE, prior="uniform")
#beta.ash.vb.null = ash(ss$betahat, ss$betasd,auto=TRUE, VB=TRUE, prior=NULL)

@@

Now, we find the fitted values for $\pi$ are mostly near 0, or at 2 (the alternative value).
```{r}
beta.ash$fitted.f
beta.ash.auto$fitted.f
#beta.ash.vb.uniform$fitted.f
#beta.ash.vb.null$fitted.f
```

@@

[for testing: note that the PosteriorMean and PositiveProb don't depend much on sigmaa vec used ]
```{r}
plot(beta.ash.auto$PosteriorMean, beta.ash$PosteriorMean, xlab="Shrunk estimate from auto method", ylab="Shrunk estimate from fixed method")
abline(a=0,b=1)
plot(beta.ash.auto$localfdr, beta.ash$localfdr, xlab="lfdr from auto method", ylab="ldfr from fixed method")
abline(a=0,b=1)
```

Also, we can see the effects of shrinkage: small estimates of $\hat\beta$ are
shrunk to close to 0. Large estimates of $\hat\beta$ are shrunk less strongly because ash recognizes that these larger $\hat\beta$ are likely
from the alternative, rather than the null.
```{r}
plot(ss$betahat,beta.ash$PosteriorMean,xlab="Observed betahat", ylab="Estimated beta (posterior mean)",ylim=c(-7,4),xlim=c(-7,4))
abline(h=0)
abline(a=0,b=1,col=2)
```

@@

Let's try: 
```{r}
beta.ash.pm = ash(ss$betahat, ss$betasd, auto=TRUE, usePointMass=TRUE)
beta.ash.pm$fitted.f
beta.ash.auto$fitted.f
plot(beta.ash.auto$localfdr,beta.ash.pm$localfdr,main="comparison of ash localfdr, with and without point mass",xlab="no point mass", ylab="with point mass",xlim=c(0,1),ylim=c(0,1))
abline(a=0,b=2)
```

Our conclusion: if we simulate data with a point mass,
and we analyze it without a point mass, we may underestimate
the false sign rate by a factor of 2. Therefore, to be conservative, we might prefer to analyze the data allowing for the point mass, or, if analyzed without a point mass, multiply estimated false sign rates by 2.

@@

Here we compare ash $q$ values with those from the qvalue package. Because of the issue above, we multiply the ash  $q$ values by 2.
```{r}
plot(qval$q,2*beta.ash$qval,main="comparison of ash and q value qvalues",xlab="qvalue", ylab="twice ash q values")
abline(a=0,b=1)
```

In this example we see that qval overestimates the actual FDR. (This
is because it assumes all the $p$ values near 1 are null, when they are not.)
```{r}
o = order(beta.ash$qval)
plot(cumsum(ss$null[o])/(1:10000),qval$qval[o],col=2,type="l", xlab="actual FDR", ylab="q value", main="Comparison of actual FDR with q value")
lines(cumsum(ss$null[o])/(1:10000),2*beta.ash$qval[o])
abline(a=0,b=1)
```

@@

plot(betahat,beta.ash$PosteriorMean,xlab="Observed betahat", ylab="Estimated beta (posterior mean)",ylim=c(-7,4),xlim=c(-7,4))
abline(h=0)
abline(a=0,b=1,col=2)

plot(qval$q,2*beta.ash$qval,main="comparison of ash and q value qvalues")
abline(a=0,b=1)

@@

plot(cumsum(truenull[o])/(1:10000),qval$qval[o],col=2,type="l")
lines(cumsum(truenull[o])/(1:10000),2*beta.ash$qval[o])
abline(a=0,b=1)
```

It seems that in this case the ash q values underestimate the
FDR slightly. Possibly this is the assymetry.
Try shrinking positive and negatives separately:
```{r}
pos= betahat>0
betapos.ash=ash(betahat[pos],s[pos])
betaneg.ash = ash(betahat[!pos],s[!pos])
lfdr = rep(0,length(betahat))
lfdr[pos] = betapos.ash$localfdr
lfdr[!pos] = betaneg.ash$localfdr
qv = qval.from.localfdr(lfdr)
o = order(qv)
plot(cumsum(truenull[o])/(1:10000),2*qv[o],type="l")
abline(a=0,b=1)
```
No, that's not it.

@@

beta.ash=ash(betahat,s)
#compute the usual zscore and corresponding p value
zscore = betahat/s
pval = pchisq(zscore^2,df=1,lower.tail=F)
qval = qvalue(pval)
plot(qval$q,2*beta.ash$qval,main="comparison of ash and q value qvalues")
abline(a=0,b=1)

@@

plot(cumsum(truenull[o])/(1:10000),qval$qval[o],col=2,type="l")
lines(cumsum(truenull[o])/(1:10000),2*beta.ash$qval[o])
abline(a=0,b=1)

>>>>

Note that $\text{lfsr}_j \geq \text{lfdr}_j$ 
because both the events $\beta_j \geq 0$
and $\beta_j \leq 0$ include the event $\beta_j=0$.
Thus, lfsr gives an upper bound for the lfdr,
and so can be used as a conservative estimate of the lfdr
if an lfdr is desired. (This may be helpful when comparing
with methods that compute an lfdr.)
In addition, we intuitively expect the difference $\text{lfsr}_j - \text{lfdr}_j$ to be small, at least when $\text{lfdr}_j$ is small, because when we are confident that a particular effect is non-zero we would expect to be similarly confident about the sign of the effect.

@@

beta.ash=ash(ss$betahat,ss$betasd)
beta.ash.auto = ash(ss$betahat, ss$betasd,auto=TRUE)
#these to test the VB version 
beta.ash.vb.uniform = ash(ss$betahat, ss$betasd,auto=TRUE, VB=TRUE, prior="uniform")
beta.ash.vb.null = ash(ss$betahat, ss$betasd,auto=TRUE, VB=TRUE, prior=NULL)

@@

Now, we find the fitted values for $\pi$ are mostly near 0, or at 2 (the alternative value).
```{r}
beta.ash$fitted.f
beta.ash.auto$fitted.f
beta.ash.vb.uniform$fitted.f
beta.ash.vb.null$fitted.f
```

@@

[for testing: note that the PosteriorMean and PositiveProb don't depend much on sigmaa vec used ]

```{r}
plot(beta.ash.auto$PosteriorMean, beta.ash$PosteriorMean, xlab="Shrunk estimate from auto method", ylab="Shrunk estimate from fixed method")
abline(a=0,b=1)
plot(beta.ash.auto$localfdr, beta.ash$localfdr, xlab="lfdr from auto method", ylab="ldfr from fixed method")
abline(a=0,b=1)

[And VB method produces similar results to EM method]
```{r}
plot(beta.ash.auto$PosteriorMean, beta.ash.vb.uniform$PosteriorMean, xlab="Shrunk estimate from auto method", ylab="Shrunk estimate from vb method")
points(beta.ash.auto$PosteriorMean, beta.ash.vb.null$PosteriorMean, col=2)
abline(a=0,b=1)
plot(beta.ash.auto$localfdr, beta.ash.vb.uniform$localfdr, xlab="lfdr from auto method", ylab="ldfr from vb method")
points(beta.ash.auto$localfdr, beta.ash.vb.null$localfdr, col=2)
abline(a=0,b=1)

Also, we can see the effects of shrinkage: small estimates of $\hat\beta$ are
shrunk to close to 0. Large estimates of $\hat\beta$ are shrunk less strongly because ash recognizes that these larger $\hat\beta$ are likely
from the alternative, rather than the null.
```{r}
plot(ss$betahat,beta.ash$PosteriorMean,xlab="Observed betahat", ylab="Estimated beta (posterior mean)",ylim=c(-7,7),xlim=c(-7,7))
abline(h=0)
abline(a=0,b=1,col=2)
```

@@

Let's try: 
```{r}
beta.ash.pm = ash(ss$betahat, ss$betasd, auto=TRUE, usePointMass=TRUE)
beta.ash.pm$fitted.f
beta.ash.auto$fitted.f
plot(beta.ash.auto$localfsr,beta.ash.pm$localfsr,main="comparison of ash localfsr, with and without point mass",xlab="no point mass", ylab="with point mass",xlim=c(0,1),ylim=c(0,1))
abline(a=0,b=1)
abline(a=0,b=2)
```

Our conclusion: if we simulate data with a point mass,
and we analyze it without a point mass, we may underestimate
the lfsr by a factor of 2. Therefore, to be conservative, we might prefer to analyze the data allowing for the point mass, or, if analyzed without a point mass, multiply estimated false sign rates by 2. In fact the latter might be preferable: even if we analyze the data with a point mass, there is going to be some unidentifiability
that means estimating the pi value on the point mass will be somewhat unreliable, and we still might underestimate the false sign rate if we rely on that estimate.

@@

Here we compare ash $q$ values with those from the qvalue package. 
```{r}
plot(qval$q,beta.ash$qval,main="comparison of ash and q value qvalues",xlab="qvalue", ylab="ash q values")
abline(a=0,b=1)
```

In this example we see that qval overestimates the actual FDR. (This
is because it assumes all the $p$ values near 1 are null, when they are not.)
```{r}
o = order(beta.ash$qval)
plot(cumsum(ss$null[o])/(1:10000),qval$qval[o],col=2,type="l", xlab="actual FDR", ylab="q value", main="Comparison of actual FDR with q value")
lines(cumsum(ss$null[o])/(1:10000),beta.ash$qval[o])
abline(a=0,b=1)
```

@@

plot(betahat,beta.ash$PosteriorMean,xlab="Observed betahat", ylab="Estimated beta (posterior mean)",ylim=c(-7,7),xlim=c(-7,7))
abline(h=0)
abline(a=0,b=1,col=2)

plot(qval$q,beta.ash$qval,main="comparison of ash and q value qvalues")
abline(a=0,b=1)

@@

plot(cumsum(truenull[o])/(1:10000),qval$qval[o],col=2,type="l")
lines(cumsum(truenull[o])/(1:10000),beta.ash$qval[o])
abline(a=0,b=1)
```

It seems that in this case the ash q values underestimate the
FDR slightly. Possibly this is the assymetry.
Try shrinking positive and negatives separately:
```{r}
pos= betahat>0
betapos.ash=ash(betahat[pos],s[pos])
betaneg.ash = ash(betahat[!pos],s[!pos])
lfdr = rep(0,length(betahat))
lfdr[pos] = betapos.ash$localfdr
lfdr[!pos] = betaneg.ash$localfdr
qv = qval.from.localfdr(lfdr)
o = order(qv)
plot(cumsum(truenull[o])/(1:10000),qv[o],type="l")
abline(a=0,b=1)
```
No, that's not it.

@@

beta.ash=ash(betahat,s)
#compute the usual zscore and corresponding p value
zscore = betahat/s
pval = pchisq(zscore^2,df=1,lower.tail=F)
qval = qvalue(pval)
plot(qval$q,beta.ash$qval,main="comparison of ash and q value qvalues")
abline(a=0,b=1)

@@

plot(cumsum(truenull[o])/(1:10000),qval$qval[o],col=2,type="l")
lines(cumsum(truenull[o])/(1:10000),beta.ash$qval[o])
abline(a=0,b=1)
